Delivering reliable software in the European scientific arena: evolution of methodologies and new trend analysis

### 2017-11-30 paper-ieee-trans.tex
add text and new image for automatic configuration

\subsection{DevOps practices}

@@

The requirements are applied at early stages in the software development
process, integrated in the Jenkins CI service \cite{indigo-jenkins}, so that any
bug or design issue is likely to be detected and corrected promptly. The source
code is publicly available in GitHub repositories under an organization called
{\sl indigo-dc} \cite{indigo-github} to increase the visibility of the product
catalogue, promoting the external contributions and software adoption. Furthermore,
the source code is compliant with community de--facto style standards,
selected from the variety of programming languages being used.

@@

Tests are triggered in each build to measure the unit and functional testing
coverage, with a recommended value of 70\% of coverage in the former case. The
documentation of the products is treated as source code, using a markup
language, automatically rendered and uploaded in online repositories
\cite{indigo-gitbook}. Changes in both documentation and source code are
human--reviewed as the last step before merging them into the production
repository. Last but not least, as already seen in the DevOps section, in order
to facilitate the usage of the INDIGO--DataCloud services, they can be deployed
automatically using either Ansible \cite{indigo-ansible} or Puppet
\cite{indigo-puppet} open--source tools.

The evaluation of the software quality is performed by measuring the values of
the metrics and Key Performance Indicators (KPIs) defined based upon the
ISO/IEC 9126 standard. These metrics cover the development, release and
maintenance phases of the software lifecycle. Development and release metrics
are obtained automatically from several sources, such as GitHub and Jenkins CI
service, and graphically displayed as GitHub pages using GrimoireLab framework
\cite{grimoirelab}. Maintenance and user support metrics are collected from the
different sources of data like GGUS \cite{ggus} and Github Issues.

@@

%\appendices
%\section{Proof of the First Zonklar Equation}
%Appendix one text goes here.
%
%% you can choose not to have a title for an appendix
%% if you want by leaving the argument blank
%\section{}
%Appendix two text goes here.

% use section* for acknowledgment
\section*{Acknowledgment}

@@

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

@@

% biography section
%
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

@@

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

>>>>

\subsection{DevOps practices}
\label{sec:devops}

@@

The requirements are applied at early stages in the software development
process, integrated in the Jenkins CI service \cite{indigo-jenkins}, so that any
bug or design issue is likely to be detected and corrected promptly. The source
code is publicly available in GitHub repositories under an organization called
{\sl indigo-dc} \cite{indigo-github} to increase the visibility of the product
catalogue, promoting the external contributions and software adoption. Furthermore,
the source code is compliant with community de--facto style standards,
selected from the variety of programming languages being used as seen in
Fig. \ref{fig:fig_codestyle}.

@@

Unit testing coverage has increased between the first ({\sl INDIGO-1}) and the
second ({\sl INDIGO-2}) INDIGO--DataCloud releases, reaching an average value of
66.69\% on the {\sl INDIGO-2} release, very close to the recommended 70\% threshold
defined by the project’s SQA base criteria \cite{indigo-d31}. A comparison graph
between the code coverage for the products involved both in {\sl INDIGO-1} and
{\sl INDIGO-2} releases is shown in Fig. \ref{fig:fig_unittest}. As it can be seen,
a very small set of products with low coverage values highly impacted the overall
result, although a significant part of these products are upstream contributions
or products with codebases not fully under the control of the project. Nevertheless,
60\% of the products were over this threshold while 80\% of them have values above
the 50\% coverage.

The documentation of the products is treated as source code, using a markup
language, automatically rendered and uploaded to online repositories
\cite{indigo-gitbook}. Changes in both documentation and source code are
human--reviewed as the last step before merging them into the production
repository.

Last but not least, as described in section \ref{sec:devops}, in order
to facilitate the usage of the INDIGO--DataCloud services, they can be deployed
automatically using either Ansible \cite{indigo-ansible} or Puppet
\cite{indigo-puppet} open--source tools. The adoption of such Configuration Management
tools not only eases the consumption of the products developed by external users or
infrastructures, but also tackles the correct installation and configuration of the
services and applications. A representative example is the number of roles currently
hosted in the Ansible Galaxy portal for the project.

Fig. \ref{fig:fig_confman} shows the number of products that offer an automated means
for deployment. It shows an increase after the {\sl INDIGO-1} release with specially
in the previous weeks of each of the two major releases.

The evaluation of the software quality is performed by measuring the values of
the metrics and Key Performance Indicators (KPIs) defined based upon the
ISO/IEC 9126 standard. These metrics cover the development, release and
maintenance phases of the software lifecycle. Development and release metrics
are obtained automatically from several sources, such as GitHub and Jenkins CI
service, and graphically displayed as GitHub pages using GrimoireLab framework
\cite{grimoirelab}. Maintenance and user support metrics are collected from the
different sources of data such as GGUS \cite{ggus} and Github Issues.

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{images/confman.png}
\caption{Trend line showing the adoption of Configuration Management tools throughout the project lifetime.}
\label{fig:fig_confman}
\end{figure*}

@@

\section*{Acknowledgment}

@@

\begin{thebibliography}{1}

@@

@@

\end{document}
